---
title: 'Statistical Inference: Peer Assessment, Part 2'
author: "S C NG"
date: "20 June, 2015"
output: pdf_document
---

## Overview
In the second portion of the assignment, we're going to analyze the ToothGrowth data in the R datasets package. We will use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose, and make the conclusion with assumptions needed.

## Setup
First of all, the following default settings and libraries are loaded
```{r}
#preset default options for Rmd, codes not shown in report
require(knitr)
opts_chunk$set(cache=TRUE, echo=TRUE)

#load required libraries for data analysis
require(ggplot2)
require(datasets)
```

## Exploratory of Data
According to the help page, the ToothGrowth dataset recorded the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).
Below R code shows help page and the data structure of ToothGrowth dataset
```{r}
?ToothGrowth
str(ToothGrowth)
```
The dataset consists of 60 observations on 3 variables: 
1. len  (numeric) = Tooth length
2. supp (factor) = Supplement type (VC or OJ).
3. dose (numeric) = Dose in milligrams.

Below boxplot summarize the distribution of Tooth length values per each three dose levels, 1 plot per each Supplement type
```{r}
g2 <- ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) 
g2 <- g2 + geom_boxplot()
g2 <- g2 + facet_grid(. ~ supp) 
g2 <- g2 + xlab("Dose in miligrams") + ylab("Tooth length") 
g2 <- g2 + guides(fill = guide_legend(title ="Supplement type"))
 g2 <- g2 + ggtitle("Boxplot for Toothgrowth dataset")
g2
```

## Data Analysis
The following section tried to use confidence intervals and hypothesis tests to compare the tooth growth by supplement type and dose

We will start with comparing the tooth length by supplement type through the following test in R
```{r}
t1 <- t.test(len~supp, paired = FALSE, var.equal = TRUE, data=ToothGrowth)
t1
```
TODO: Analysis???

Then we will check the tooth lenght by dose
TODO: Analysis between len and dose???


##Assumptions
For the above tests, we assume the len values in the dataset are normally distributed